---
output:
  md_document:
      variant: gfm
html_preview: true
---

# Extracting data from google's COVID-19 community mobility reports

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Google has recently published [reports on the mobilty data](https://www.google.com/covid19/mobility/) for different contires/regions. However, I could not find the raw data behind the plots. Here is my attemp to recover the raw data.

### Downloading the report and extract the data graphs

This is done using [shell commands](getdata.md).

### Process one example image in R (Grocery & pharmacy in Singapore)

Load libraries and an example image

```{r message=FALSE, warning=FALSE}
library(grid)
library(dplyr)
library(png)
library(ggplot2)
img <- readPNG("2020-03-29_SG_Mobility_Report_en.2.png")
```


We can extract the color with a digital color meter:

 - Graph color rgb(81, 135,237) or rgb(0.32, 0.53, 0.93)
 - Grid color rgb(216, 218, 222) or rgb(0.85, 0.85, 0.87)


Find the graph
```{r}
delta <- 0.08

graph.sel <- abs(img[,,1] - 0.32) < delta & abs(img[,,2] - 0.53) < delta & abs(img[,,3] - 0.93) < delta

## Validate the selection
img[,,1][graph.sel] <- 1
img[,,2][graph.sel] <- 0
img[,,3][graph.sel] <- 0

grid.raster(img[,,])
```

Find the grid
```{r}
delta <- 0.08
grid.sel <- abs(img[,,1] - 0.85) < delta & abs(img[,,2] - 0.85) < delta & abs(img[,,3] - 0.85) < delta

img[,,1][grid.sel] <- 0
img[,,2][grid.sel] <- 1
img[,,3][grid.sel] <- 0

grid.raster(img[,,])
```

Find y values
```{r}
gridlines.y <- data.frame(rows=which(grid.sel[, round(dim(grid.sel)[2]/2)])) %>% 
    mutate(diff=rows-lag(rows)) %>% 
    filter(diff >2 | is.na(diff)) %>% 
    pull(rows)
zero <- (gridlines.y[1] + last(gridlines.y))/2
scale <- 1.6/abs(gridlines.y[1] - last(gridlines.y))

y <- (zero-apply(graph.sel,2,function(x) mean(which(x))) )* scale

# check extracted data shape
plot(y)
```

Find x values
```{r}
## find position of the ticks
tmp <- grid.sel[last(gridlines.y):(last(gridlines.y) + 10),] %>% colSums()

ticks.x <- data.frame(cols=which(tmp > 5)) %>% 
    mutate(diff=cols-lag(cols)) %>% 
    filter(diff >2 | is.na(diff)) %>% pull(cols) 

step <- (last(ticks.x)-ticks.x[1])/42## 43 days
x.sel <- round(seq(ticks.x[1] +1, length=43, by=step))

extracted <- data.frame(Date=seq(as.Date("2020/2/16"), as.Date("2020/3/29"), by=1),
           Relative_mobility=y[x.sel])

## You can now save this as a csv or plain text
head(extracted)
```

Let's plot
```{r}
ggplot(extracted, aes(x=Date, y=Relative_mobility*100)) + 
    geom_area(fill='grey') + 
    geom_line(lwd=2) + 
    theme_bw() 
```

Session information
```{r}
sessionInfo()
```
